---
title: "Data624 Time Series Graphics Assignment1"
author: "Ritesh Lohiya"
date: "February 7, 2019"
output: html_document
---

#Data624 Time Series Graphics Assignment1

#Chapter 2:


```{r}
suppressMessages(suppressWarnings(library(fpp2)))
suppressMessages(suppressWarnings(library(readxl)))
```

2.1. Use the help function to explore what the series gold, woolyrnq and gas represent.



```{r}
str(gold)
str(woolyrnq)
str(gas)

# a. Use autoplot() to plot each of these in separate plots.

autoplot(gold)
autoplot(woolyrnq)
autoplot(gas)


# b. What is the frequency of each series? Hint: apply the frequency() function.

frequency(gold)
frequency(woolyrnq)
frequency(gas)

# c. Use which.max() to spot the outlier in the gold series. Which observation was it?

which.max(gold)
gold[which.max(gold)]

```

2.2. Download the file tute1.csv from the book website, open it in Excel (or some other spreadsheet application), and review its contents. You should find four columns of information. Columns B through D each contain a quarterly series, labelled Sales, AdBudget and GDP. Sales contains the quarterly sales for a small company over the period 1981-2005. AdBudget is the advertising budget and GDP is the gross domestic product. All series have been adjusted for inflation.

```{r}
# a. You can read the data into R with the following script:

tute1 <- read.csv("C:/Users/rites/Documents/GitHub/Data624_Assignment1/tute1.csv", header=TRUE)
View(tute1)

# b. Convert the data to time series

mytimeseries <- ts(tute1[,-1], start=1981, frequency=4)

# (The [,-1] removes the first column which contains the quarters as we don't need them now.)

# c. Construct time series plots of each of the three series

autoplot(mytimeseries, facets=TRUE)

# Check what happens when you don't include facets=TRUE.

autoplot(mytimeseries)

#The labels are seperated and colored 

```

2.3. Download some monthly Australian retail data from the book website. These represent retail sales in various categories for different Australian states, and are stored in a MS-Excel file.

```{r}
# a. You can read the data into R with the following script:

retaildata <- readxl::read_excel("C:/Users/rites/Documents/GitHub/Data624_Assignment1/retail.xlsx", skip=1)

# The second argument (skip=1) is required because the Excel sheet has two header rows.

# b. Select one of the time series as follows (but replace the column name with your own chosen column):

myts <- ts(retaildata[,"A3349873A"],frequency=12, start=c(1982,4))

#c. Explore your chosen retail time series using the following functions:

#Can you spot any seasonality, cyclicity and trend? What do you learn about the series?

autoplot(myts)
ggseasonplot(myts)
ggsubseriesplot(myts)
gglagplot(myts, lags = 12)
ggAcf(myts)

# WE can see seasonality and trend here
```

2.6. Use the following graphics functions: autoplot(), ggseasonplot(), ggsubseriesplot(), gglagplot(), ggAcf() and explore features from the following time series: hsales, usdeaths, bricksq, sunspotarea, gasoline.

```{r}
# Can you spot any seasonality, cyclicity and trend?

# What do you learn about the series?

# hsales:

autoplot(hsales)
ggseasonplot(hsales)
ggsubseriesplot(hsales)
gglagplot(hsales)
ggAcf(hsales, lag.max = 400)

# We can see seasonality and cyclicity. The cycle period is ~100 months.

# usdeaths:

autoplot(usdeaths)
ggseasonplot(usdeaths)
ggsubseriesplot(usdeaths)
gglagplot(usdeaths)
ggAcf(usdeaths, lag.max = 60)

# We can see seasonality.

# bricksq:

autoplot(bricksq)
ggseasonplot(bricksq)
ggsubseriesplot(bricksq)
gglagplot(bricksq)
ggAcf(bricksq, lag.max = 200)

# We can see slight seasonality and very good trend.

# sunspotarea:

autoplot(sunspotarea)
# ggseasonplot(sunspotarea) 

# Data are not seasonal

# ggsubseriesplot(sunspotarea)

# Data are not seasonal

gglagplot(sunspotarea)
ggAcf(sunspotarea, lag.max = 50)

# We can see strong cyclicity.

# gasoline:

autoplot(gasoline)
ggseasonplot(gasoline)

#ggsubseriesplot(gasoline)

# Each season requires at least 2 observations. That is not present

gglagplot(gasoline)
ggAcf(gasoline, lag.max = 1000)

# We can see seasonality and trend.
```








